### 2024-07-10 Rmd/sample_scale_vs_scale_sample.Rmd

---
title: "Sample scale vs scale sample"
author: "Brett Stacy"
date: "2024-06-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

# Can we scale a boostrap sample?

Can we take a boostrap sample of hauls, then scale it and it will equal the boostrapped scale?

## Bootstrap then scale

Bootstrap the hauls, then apply the scaling factors to

```{r}
iters = 10
haul_size = c(100, 100, 200, 300, 200, 100)
DT = data.table(hauls = letters[1:6], lengths = c(41:46))

r_hauls = purrr::map(1:iters, ~ sample(hauls, replace = T)) # resampled hauls

```

## Bootstrap the scale