### 2024-08-02 Rmd/compare_output_pre_and_post_adding_strata.Rmd
clean some things up - rerun with trip_strata save Steve's y2 object to use in sql_downloads add section in sql_downloads to inspect and add trip_join eventually

---
title: "Compare output pre and post adding strata"
author: "Brett Stacy"
date: "2024-07-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
```

# Data

```{r}
out1 = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/ISS_output_to_email_7_23_24.RDS")

out2 = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/ISS_1000iters_output_7_30_24.RDS")

```

# Output

## Time series of sample size

```{r}

par(mfrow = c(3,1), mar = c(2, 4.5, .5, .5))
plot(out1$iss_length$YEAR, out1$iss_length$nss, type = "l", ylab = "nominal sample size", xlab = "", cex.lab = 1.3)

plot(out1$iss_length$YEAR, out1$iss_length$nhls, type = "l", ylab = "number of hauls", xlab = "", col = "black", cex.lab = 1.3)

plot(out1$iss_length$YEAR, out1$iss_length$iss, type = "l", ylab = "ISS", xlab = "year", cex.lab = 1.3) 
lines(out2$iss_length$YEAR, out2$iss_length$iss, col = "blue")
legend("topleft", lty = c(1,1), col = c("black", "blue"), legend = c("without sampling strata", "with sampling strata"))

m1 = lm(out2$iss_length$nss ~ out2$iss_length$iss)
# summary(m1)

m2 = lm(out2$iss_length$nhls ~ out2$iss_length$iss)
# summary(m2)

```

## Boxplots of RSS

```{r}
df = data.frame(rss = out2$rss_length$rss, year = as.factor(out2$rss_length$YEAR))
# plot(df$year, df$rss, xaxt = "n", yaxt = "n", ylab = "", xlab = "")
plot(df$year, df$rss, ylab = "rss", cex.lab = 1.3)
```

## All together

```{r, fig.height=8}
# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
par(mfrow = c(4,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out1$iss_length$YEAR, out1$iss_length$nss, type = "l", ylab = "nominal sample size", xlab = "", cex.lab = 1.3, xaxt = "n")

plot(out1$iss_length$YEAR, out1$iss_length$nhls, type = "l", ylab = "number of hauls", xlab = "", col = "black", cex.lab = 1.3, xaxt = "n")

plot(df$year, df$rss, ylab = "RSS", cex.lab = 1.3, xaxt = "n")

plot(out1$iss_length$YEAR, out1$iss_length$iss, type = "l", ylab = "ISS", xlab = "year", cex.lab = 1.3) 
lines(out2$iss_length$YEAR, out2$iss_length$iss, col = "blue")
legend("topleft", lty = c(1,1), col = c("black", "blue"), legend = c("without sampling strata", "with sampling strata"))

```

```{r from my phd, eval=F}
box.plot.status.proj2 = function(proj.raw, ...){
  sta_err = list()
  for(i in 1:length(proj.raw)){
    sta_err[[i]] = SSB.err.rel.proj(output = proj.raw[[i]], final.year.ind = grep("SSB", names(proj.raw[[i]]$M.TR.))[length(grep("SSB", names(proj.raw[[i]]$M.TR.)))])
  }

  # temp.ass.names = as.character(1:length(sta_err))
  # temp.ass.names = 1:length(sta_err)
  temp.ass.names = as.character(assessment_years)
  df = data.frame(error = c(unlist(sta_err)), type = as.factor(c(sapply(temp.ass.names, function(x) rep(x, 1000)))))

  # plot(df$type, df$error, ylim = c(-0.5, .5), xaxt = "n", yaxt = "n", ylab = "", xlab = "", ...)
  plot(df$type, df$error, xaxt = "n", yaxt = "n", ylab = "", xlab = "", ...)
  abline(h = 0, lwd = 0.5, lty = 2, col = "black")

  abline(v = 13, lty = 2, lwd = 0.5)

}

par(mfrow = c(2,3), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
  for (i in c(2,4,6)) {
    box.plot.status.proj2(out.projections.raw.I$`0.6`[[i]], ylim = c(-0.35, .8))
    text(2, .75, plot.letters[i/2], cex = subplot.letter.cex)
    if(i %in% c(1,2)) axis(side = 2, las = 2)
  }
  for (i in c(2,4,6)) {
    box.plot.status.proj2(out.projections.raw.R$`0.6`[[i]], ylim = c(-0.36, .12))
    text(2, .1, plot.letters[i/2+3], cex = subplot.letter.cex)
    if(i %in% c(2,4,6)) axis(side = 1, at = c(1, 7, 13, 19), labels = c(1996, 2002, 2008, 2014))
    if(i %in% c(1,2)) axis(side = 2, las = 2)
  }
mtext(text = expression(paste("SSB"["status"], " RE")), side = 2, outer = T, line = 2, las = 0, cex = 1.2)
```