### 2024-08-19 Rmd/steve_y2_object.Rmd

---
title: "Steve y2 data"
author: "Brett Stacy"
date: "2024-08-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Steve y2 object

```{r}
load("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/Data/From Steve B/y2_data.RData")

# save as .RDS to load into run.r script
saveRDS(y2_nosex, "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/inputs/y2_nosex_ebs_pcod_Steve.RDS")

saveRDS(y2_sex, "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/inputs/y2_sex_ebs_pcod_Steve.RDS")

```