### 2024-08-27 Rmd/tidytable_left_join.Rmd
add age data frame processing. prepare for baseISS next.

---
title: "Tidy Table Practice"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Practice Tidytable left_join()

```{r}
library(tidytable)
```

Simple example from the help file. 
```{r}

df1 <- data.table(x = c("a", "a", "b", "c"), y = 1:4)
df2 <- data.table(x = c("a", "b"), z = 5:6)
df1
df2

df1 %>% left_join(df2)

```

The left_join found that the column x matched between df1 and df2. It kept all the columns in df1, but added the unshared column from df2, z, for the rows where the shared column, x, matched between df1 and df2. Where there was no match, it put an NA for column z in the new data.table.

Now let's see what happens when we add an additional shared column, w.

```{r}
df1 <- data.table(x = c("a", "a", "b", "c"), w = c("a", "a", "b", "c"), y = 1:4)
df2 <- data.table(x = c("a", "b"), w = c("a", "c"), z = 5:6)
df1
df2

df1 %>% left_join(df2)
```

This tries to match the row entries for columns x and w between df1 and df2. Where they match, the corresponding z value is included. Where they do not match, it puts an NA.

Now let's see what happens when there are multiple matches in the right data frame:
```{r}
df1 <- data.table(x = c("a", "a", "b", "c"), w = c("a", "a", "b", "c"), y = 1:4)
df2 <- data.table(x = c("a", "b"), w = c("a", "c"), z = 5:6)
df1
df2

df2 %>% left_join(df1)
```

It expands rows where there are multiple matches, while keeping all original row combinations of the left data frame. Where there is no match, it puts an NA for the unshared column from the right data frame.

Now what happens if there are multiple matches but the unshared column in the right data frame has two of the same exact row entries
```{r}
df1 <- data.table(x = c("a", "a", "b", "c"), w = c("a", "a", "b", "c"), y = c(1, 1, 3, 4))
df2 <- data.table(x = c("a", "b"), w = c("a", "c"), z = 5:6)
df1
df2

df2 %>% left_join(df1)
```
It does the same thing as before. It expands the rows by the number of matches.