### 2024-09-09 Rmd/post_stratify_sex_weighting_question.Rmd
get rid of deprecated length only scripts. Work on incorporating non-fake age resampling functionality. Initiate documenting WEIGHT1 questions for age.

---
title: "Should I re-calculate WEIGHT1 when post-stratifying by sex?"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidytable)
```

# Description
Sex is the only inter-haul characteristic (in length data) that can be post-stratified by. But does this mean that WEIGHT1 needs to be recalculated when sex is filtered by to post-stratify by it? WEIGHT1 = SUM_FREQUENCY/YAGMH_SFREQ. i.e., the number of fish of a particular length divided by the total number of fish lengthed in a haul.

Load a sex-differentiated input data frame:
```{r}
lfreq_data = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/inputs/y2_sex_ebs_pcod_Steve_TRIP_STRATA.RDS")

```

Determine if SUM_FREQUENCY and YAGMH_SFREQ are sex-dependent. 
```{r}
lfreq_data[HAUL_JOIN=="H101530"] %>% 
  tidytable::glimpse()

lfreq_data[HAUL_JOIN=="H101530", .(SEX, LENGTH, YAGMH_SFREQ, SUM_FREQUENCY, WEIGHT1)]

lfreq_data[HAUL_JOIN=="H101530", sum(WEIGHT1)]

lfreq_data[HAUL_JOIN=="H101530", .(SEX, LENGTH, YAGMH_SFREQ, SUM_FREQUENCY, WEIGHT1, WEIGHT2, WEIGHT4)]

```

# Conclusion
Sex can only be post-stratified by if it is included as a column in the input data frame. In the case of EBS Pcod, the WEIGHT1 column is calculated as combined-sex (see above r code). So there are two important questions:

1. When a single sex is filtered, can we use the existing WEIGHT1 values?
2. When WEIGHT1 is recalculated in fishery_props.R, is using the existing YAGMH_SFREQ value appropriate? Because it will often be larger than if we recalculated it.