### 2024-09-09 README.Rmd
update documentation

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# baseISS

<!-- badges: start -->
<!-- badges: end -->

The goal of baseISS is to develop a preliminary method for calculating input sample size (ISS) for fishery-dependent composition data. This data includes length- and age-frequency data collected by the NORPAC observer program and excludes fishery-independent data from surveys. Fishery-dependent composition data represents an important element of the 26 statistical catch at age stock assessment models for Alaskan fisheries managed by the NOAA AFSC. The ISS determines the precision attached to this data type, and therefore regulates its leverage in assessment model outputs such as biomass predictions.

The structure and functionality of this package resembles surveyISS with some notable differences that accomodate the neuances of fishery-dependent data. These are covered in the next section, "Important Notes".

## Important Notes

### Input Data
The following conditions of an input data frame for any species must be met:
1. (length data) There exists a WEIGHT1 column that represents the haul-level proportions-at-length of the observed lengths. This is calculated by dividing the frequency of the observed length within a haul by the total number of fish lengthed in that haul. It should be a standard calculation that is exactly the same for every species.

### Current Functionality

### Developing Functionality

## Installation

You can install the development version of baseISS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Brett-Stacy/baseISS")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(baseISS)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.