### 2024-09-17 README.Rmd
investigate sample size for adding change_N functionality. Initiate test run script for adding functionality.

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# baseISS

<!-- badges: start -->
<!-- badges: end -->

The goal of baseISS is to develop a preliminary method for calculating input sample size (ISS) for fishery-dependent composition data. This data includes length- and age-frequency data collected by the NORPAC observer program and excludes fishery-independent data from surveys. Fishery-dependent composition data represents an important element of the 26 statistical catch at age stock assessment models for Alaskan fisheries managed by the NOAA AFSC. The ISS determines the precision attached to this data type, and therefore regulates its leverage in assessment model outputs such as biomass predictions.

The structure and functionality of this package resembles surveyISS with some notable differences that accommodate the nuances of fishery-dependent data. These are covered in the next section, "Important Notes".

## Important Notes

### Input Data
The following conditions of an input data frame for any species must be met:
1. (length data) There exists a WEIGHT1 column that represents the haul-level proportions-at-length of the observed lengths. This is calculated by dividing the frequency of the observed length within a haul by the total number of fish lengthed in that haul. It should be a standard calculation that is exactly the same for every species.

2. Length data be condensed with sum_frequency column.

3. Age data include associated lengths.

4. Age data flattened, no "sum_frequency" column.

5. For EBS Pcod, age data has only been used when there exists alredy pre-processed length data associated with it. I.e., I matched the y2 length data object recieved from Steve with age data I downloaded when testing the baseISS functionality on age data. This was necessary becasue I needed quantities like YAGM_SNUM to perform the weighted expansion.

### Current Functionality

1. Length ISS must be a separate analysis to age ISS. Unlike surveyISS, this is necessary for several reasons: a. there will be many more hauls with lengths than ages, b. it is likely that the user options such as post_stratify will be different between length data and age data, c. NOTE ANOTHER ONE WHEN YOU REMEMBER.

2. You can't chose the species from a big data set containing all the species. This is because data sets for each species should include unique columns necessary to apply their unique expansion methods. For EBS Pcod for example, these columns are only available after much pre-processing and calculating of the raw data.

### Developing Functionality

- subsampling

## Installation

You can install the development version of baseISS from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("Brett-Stacy/baseISS")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(baseISS)
## basic example code
```

What is special about using `README.Rmd` instead of just `README.md`? You can include R chunks like so:

```{r cars}
summary(cars)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.