### 2024-09-17 Rmd/haul_sample_size_investigation.Rmd

---
title: "Inspect YAGMH_SFREQ"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidytable)
library(ggplot2)
```

```{r}
lfreq_data = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/inputs/y2_sex_ebs_pcod_Steve_TRIP_STRATA.RDS")

```

# Introduction
We want to see how ISS changes when sample size is reduced or increased. The target sample size per haul has been 20 fish lengthed for some time now. Is that too many? Is that too few? Can baseISS be used as a framework to optimize sample size? Additionally, do adjustments to target sample size impact ISS differently when expansion method changes? I.e., is a change to sample size robust to expansion method applied? -Or- At what sample size does the expansion method not matter? Are we going to test this just for one or a couple years? Or for the whole time series?

## Inspect the distribution of a recent length composition data set
```{r}
lfreq_data[YEAR==2022, ] %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) %>% 
  tidytable::select(YAGMH_SFREQ) -> temp

hist(temp$YAGMH_SFREQ)
summary(temp$YAGMH_SFREQ)

```
Looks pretty normal

## Average sample size 
```{r}
lfreq_data %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) %>% 
  tidytable::summarize(avg_sample_N = mean(YAGMH_SFREQ), .by = YEAR) -> temp

plot(temp$YEAR, temp$avg_sample_N, type = "l")

plot(temp$YEAR[23:47], temp$avg_sample_N[23:47], type = "l")

```
Sample size varies quite a bit until 1999 when it levels out to about 20.

## Are they normal?
Are the ~20 sample size years normally distributed?

plot the boxplots over time

```{r}
lfreq_data %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) -> temp
df = data.frame(sampl_N = temp$YAGMH_SFREQ, year = as.factor(temp$YEAR))
plot(df$year, df$sampl_N, ylab = "Haul Sample N", cex.lab = 1.3)

lfreq_data[YEAR %in% 1999:2023,] %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) -> temp
df = data.frame(sampl_N = temp$YAGMH_SFREQ, year = as.factor(temp$YEAR))
plot(df$year, df$sampl_N, ylab = "Haul Sample N", cex.lab = 1.3)
plot(df$year, df$sampl_N, ylab = "Haul Sample N", cex.lab = 1.3, ylim = c(0, 100))
plot(df$year, df$sampl_N, ylab = "Haul Sample N", cex.lab = 1.3, ylim = c(18, 22))

temp %>% 
  tidytable::mutate(YEAR = factor(YEAR)) %>% 
  ggplot(aes(y=YAGMH_SFREQ, x=YEAR)) + 
  geom_violin() +
  ylim(18, 23)

# recent years wisker plot
temp %>% 
  tidytable::mutate(YEAR = factor(YEAR)) %>% 
  tidytable::mutate(sd_YAGMH_SFREQ = sd(YAGMH_SFREQ), .by = YEAR) %>% 
  tidytable::summarise(mean_YAGMH_SFREQ = mean(YAGMH_SFREQ), sd_YAGMH_SFREQ = unique(sd_YAGMH_SFREQ), .by = YEAR) %>% 
  ggplot(aes(x = YEAR, y = mean_YAGMH_SFREQ)) +
  geom_line(aes(group = 1)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean_YAGMH_SFREQ - sd_YAGMH_SFREQ, ymax = mean_YAGMH_SFREQ + sd_YAGMH_SFREQ), width = .2) +
  labs(y = "Haul Sample N (mean +- sd)", x = "Year")

# all years wisker plot mean
lfreq_data %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) -> temp
temp %>% 
  tidytable::mutate(YEAR = factor(YEAR)) %>% 
  tidytable::mutate(sd_YAGMH_SFREQ = sd(YAGMH_SFREQ), .by = YEAR) %>% 
  tidytable::summarise(mean_YAGMH_SFREQ = mean(YAGMH_SFREQ), sd_YAGMH_SFREQ = unique(sd_YAGMH_SFREQ), .by = YEAR) %>% 
  ggplot(aes(x = YEAR, y = mean_YAGMH_SFREQ)) +
  geom_line(aes(group = 1)) +
  geom_point(size = 4) +
  geom_errorbar(aes(ymin = mean_YAGMH_SFREQ - sd_YAGMH_SFREQ, ymax = mean_YAGMH_SFREQ + sd_YAGMH_SFREQ), width = .2) +
  labs(y = "Haul Sample N (mean +- sd)", x = "Year")

# all years wisker plot mean + median
lfreq_data %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) -> temp
temp %>% 
  tidytable::mutate(sd_YAGMH_SFREQ = sd(YAGMH_SFREQ), .by = YEAR) %>% 
  tidytable::summarise(mean_YAGMH_SFREQ = mean(YAGMH_SFREQ), sd_YAGMH_SFREQ = unique(sd_YAGMH_SFREQ), median_YAGMH_SFREQ = median(YAGMH_SFREQ), .by = YEAR) %>% 
  ggplot() +
  geom_line(aes(x = YEAR, y = mean_YAGMH_SFREQ), color = "blue") +
  geom_point(aes(x = YEAR, y = mean_YAGMH_SFREQ), size = 4) +
  geom_errorbar(aes(x = YEAR, y = mean_YAGMH_SFREQ, ymin = mean_YAGMH_SFREQ - sd_YAGMH_SFREQ, ymax = mean_YAGMH_SFREQ + sd_YAGMH_SFREQ), width = .2) +
  labs(y = "Haul Sample N (mean +- sd)", x = "Year") +
  geom_line(aes(x = YEAR, y = median_YAGMH_SFREQ), color = "red")

# recent years wisker plot mean + median
lfreq_data[YEAR %in% 1999:2023,] %>% 
  tidytable::distinct(HAUL_JOIN, .keep_all = TRUE) -> temp
temp %>% 
  tidytable::mutate(sd_YAGMH_SFREQ = sd(YAGMH_SFREQ), .by = YEAR) %>% 
  tidytable::summarise(mean_YAGMH_SFREQ = mean(YAGMH_SFREQ), sd_YAGMH_SFREQ = unique(sd_YAGMH_SFREQ), median_YAGMH_SFREQ = median(YAGMH_SFREQ), .by = YEAR) %>% 
  ggplot() +
  geom_line(aes(x = YEAR, y = mean_YAGMH_SFREQ), color = "blue") +
  geom_point(aes(x = YEAR, y = mean_YAGMH_SFREQ), size = 4) +
  geom_errorbar(aes(x = YEAR, y = mean_YAGMH_SFREQ, ymin = mean_YAGMH_SFREQ - sd_YAGMH_SFREQ, ymax = mean_YAGMH_SFREQ + sd_YAGMH_SFREQ), width = .2) +
  labs(y = "Haul Sample N (mean +- sd)", x = "Year") +
  geom_line(aes(x = YEAR, y = median_YAGMH_SFREQ), color = "red")

```