### 2024-09-17 Rmd/does_subsampling_WOR_matter.Rmd
change length N: establish option naming convention.

---
title: "Does it matter if we subsample WR or WOR?"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidytable)
```

# Introduction
I can't wrap my head around the justification to subsample an existing sample WOR. Does it even matter which one we do? Do a simulation to test it empirically.

More considerations:
  - if we were taking a LARGER sample, say 125%, we would sample WR right? Test this too!
  - the principle is: whatever sample we take, we want it to be representative by proportion of the original sample.

# Special Note to remember
If the optimal sample size depends on the expansion method (or post-stratification for that matter) for one species, then I think the optimal sample size will have to be species specific. Because lots of species have different expansion methods, that arguably should remain different.

## WR vs WOR subsample lower
Make up some data. Sample the same size WR. Then do two things: 1. Subsample WR, 2. subsample WOR. Do these things many times, then summarize them and compare.

```{r}

x = c(rep(1, 70), rep(0, 30))

y1 = matrix(NA, nrow = 50, ncol = 100)
y2 = matrix(NA, nrow = 50, ncol = 100)
for(i in 1:100){
  temp = sample(x, 100, replace = TRUE)
  y1[,i] = sample(temp, 50, replace = FALSE)
  y2[,i] = sample(temp, 50, replace = TRUE)
}

colMeans(y1) %>% mean()
colMeans(y2) %>% mean()

```
  I'm getting the same result either WR or WOR.

## WR vs WOR subsample higher
same as above but subsample higher

```{r, eval=F}

x = c(rep(1, 70), rep(0, 30))

y1 = matrix(NA, nrow = 150, ncol = 100)
y2 = matrix(NA, nrow = 150, ncol = 100)
for(i in 1:100){
  temp = sample(x, 100, replace = TRUE)
  y1[,i] = sample(temp, 150, replace = FALSE)
  y2[,i] = sample(temp, 150, replace = TRUE)
}

colMeans(y1) %>% mean()
colMeans(y2) %>% mean()

```
This doesn't work! Because you can't take a sample from a population larger than the population WOR. This makes sense because once you have already chosen each element of a population and not replaced it, you will run out of elements to chose from once you reach the total number in the population.

# Conclusion
Just use WR in baseISS.

BUT, now the question is, do I subsample the samples or run the whole thing again with the different N? I think running the whole thing again is better because we further randomize the hauls chosen by re-randomize the trips and hauls. But perhaps it does not matter because on average, both sets of resampled trips/hauls should be equally representative at high enough iterations. in SWO, they subsample right after boot_length for every iteration. This means they take the same resampled hauls to subsample. I'm not sure why they do this instead of resample the hauls again. Is it just for speed optimization?

For now, I'm going to code it into the existing boot_length function and necessitate running the whole thing over again if we want to resample. One potential advantage of doing it this way is that, since the sampling rate varies wildly prior to 1999, the years for which subsampling is applied will probably not be the full data year span. Therefore, the existing yrs option in fishery_iss can be used to control years to apply subsamples to. I can consider doing it the other way in the future. Start with length. start with a fixed value, 10, just for 1999 and up I guess. Note that this cannot be applied when post_strata = "SEX" because the 20 fish target is sex-independent.

Min/max problem: In SWO, the min(.N, samples) is taken. This is done instead of max() because they only considered reducing the sampling effort. I need to decide what to do with baseISS because I want to investigate greater-than sampling rates. Perhaps have a conditional statement that conditions on > or < the median YAGMH_SFREQ for the yrs spanned.

option = years, proportion/value, type, grp? SHould I code in post-strata-specific subsampling? min/max stuff?