### 2024-09-20 Rmd/Inspect_Outputs/change_sample_size_1.0.Rmd
run 100 iters of N=10, .N, 30 length per haul. Tested - works. Plot in Rmd shows results.

---
title: "Inspect Different Length Sample Size impact on ISS 1.0"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidytable)
library(ggplot2)
```

# Data Generated With:
C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS/run/run_ISS_0.5.1.R

```{r}

# speed: this took 1000 seconds
out.10 = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/haul_sample_size_change_V2_N_samples_10_ISS_100iters_output_2024-09-19.RDS")

# speed: this took 1000 seconds
out.N = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/haul_sample_size_change_V2_N_samples_.N_ISS_100iters_output_2024-09-19.RDS")

# speed: this took 1300 seconds
out.30 = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/haul_sample_size_change_V2_N_samples_30_ISS_100iters_output_2024-09-19.RDS")

```

# Introduction
Plot the ISS resulting from three levels of haul-level sample size:
1. min(.N, 10)
2. .N
3. min(.N, 30)

ancillary details:
- sample WR
- years 1999-2023 (median = 20 fish per haul)
- fixed values, not proportions

## Compare

Plot
```{r}

plot(out.10$length$iss$YEAR, out.10$length$iss$iss, type = "l", ylab = "ISS", xlab = "Year", col = "blue", ylim = c(0, 11000))
lines(out.N$length$iss$YEAR, out.N$length$iss$iss, col = "black")
lines(out.30$length$iss$YEAR, out.30$length$iss$iss, col = "red")
legend("topright", legend = c(".N", "min(.N, 10)", "min(.N, 30)"), col = c("black", "blue", "red"), lty = c(1,1,1))

```

Stats
```{r}
(out.10$length$iss$iss/out.N$length$iss$iss) %>% round(digits = 2) %>% 
  plot(type = "l")

(out.10$length$iss$iss/out.N$length$iss$iss) %>% round(digits = 2) %>% mean
```

# Conclusion
.N and 30 were equivalent. 10 was an average of 46% of .N, so about half. Implying uncertainty decreases proportional to sample size.

Try a more advanced option for changing length N where a separate sample function is called as a condition in boot_length.