### 2024-09-24 Rmd/Inspect_Outputs/change_sample_size_2.1.Rmd
run with 100 iters; N=10,.N,30.

---
title: "Inspect Different Length Sample Size impact on ISS 2.1"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(tidytable)
library(ggplot2)
```

# Data Generated With:
C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS/run/run_ISS_0.6.1.R

```{r}

# speed: this took ~1000 seconds
out.10 = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/haul_sample_size_change_V2.4_N_samples_10_ISS_100iters_output_2024-09-23.RDS")

# speed: this took ~1000 seconds
out.N = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/haul_sample_size_change_V2.4_N_samples_.N_ISS_100iters_output_2024-09-23.RDS")

# speed: this took ~1000 seconds
out.30 = readRDS("C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/haul_sample_size_change_V2.4_N_samples_30_ISS_100iters_output_2024-09-23.RDS")

```

# Introduction
Plot the ISS resulting from three levels of haul-level sample size:
1. 10
2. length(flex.vec)
3. 30

ancillary details:
- sample WR
- years 1999-2023 (median = 20 fish per haul)
- fixed values, not proportions

## Compare

Plot
```{r}

plot(out.10$length$iss$YEAR, out.10$length$iss$iss, type = "l", ylab = "ISS", xlab = "Year", col = "blue", ylim = c(0, 11000))
lines(out.N$length$iss$YEAR, out.N$length$iss$iss, col = "black")
lines(out.30$length$iss$YEAR, out.30$length$iss$iss, col = "red")
legend("topright", legend = c(".N", "10", "30"), col = c("black", "blue", "red"), lty = c(1,1,1))

```

Stats
```{r}
# 10
(out.10$length$iss$iss/out.N$length$iss$iss) %>% round(digits = 2) %>% 
  plot(type = "l")
(out.10$length$iss$iss/out.N$length$iss$iss) %>% round(digits = 2) %>% mean

# 30
(out.30$length$iss$iss/out.N$length$iss$iss) %>% round(digits = 2) %>% 
  plot(type = "l")
(out.30$length$iss$iss/out.N$length$iss$iss) %>% round(digits = 2) %>% mean

```

# Conclusion
similar to last run.