### 2024-10-15 Rmd/Inspect_Outputs/3x3_grid_sample_size_expansion_1.0.Rmd
implement max_length functionality to define the maximum length bin desired in the output. tested - works.

---
title: "Inspect 3x3 V1"
author: "Brett Stacy"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = "hide", fig.width = 8.5, fig.height = 5, dpi = 600)
library(tidytable)
library(data.table)
library(reshape2)
library(ggplot2)
library(knitr)
opts_knit$set(eval.after = "fig.cap")
```

# Introduction
3x3 grid: sample size 10, 20, 30 fixed X expansion none, haul only, haul+month. For both with and without expand by sampling strata.

## Data 
```{r}
out = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/3x3_grid_V1_ISS_200_iters_2024-10-09.RDS")

out2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/3x3_grid_V2_ISS_200_iters_2024-10-10.RDS")

year = out$samples_10$expansion_NULL$length$iss$YEAR

samples_global = c(10, 20, 30)
expansion_global = list(NULL, "haul_numbers", c("haul_numbers", "month_numbers"))
expansion_names_global = c("NULL", "haul_numbers", c("haul_month_numbers"))

```

## Plot ISS
```{r, eval=F}

df = tidytable(year = year,
                N10_Enull = out$samples_10$expansion_NULL$length$iss$iss,
                N10_Eh = out$samples_10$expansion_haul_numbers$length$iss$iss,
                N10_Ehm = out$samples_10$expansion_haul_month_numbers$length$iss$iss,
                N20_Enull = out$samples_20$expansion_NULL$length$iss$iss,
                N20_Eh = out$samples_20$expansion_haul_numbers$length$iss$iss,
                N20_Ehm = out$samples_20$expansion_haul_month_numbers$length$iss$iss,
                N30_Enull = out$samples_30$expansion_NULL$length$iss$iss,
                N30_Eh = out$samples_30$expansion_haul_numbers$length$iss$iss,
                N30_Ehm = out$samples_30$expansion_haul_month_numbers$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point()

```

## Plot ISS V2
```{r}

df = tidytable(year = year,
               N10_Enone = out$samples_10$expansion_NULL$length$iss$iss,
               N20_Enone = out$samples_20$expansion_NULL$length$iss$iss,
               N30_Enone = out$samples_30$expansion_NULL$length$iss$iss,
               N10_Emedium = out$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Emedium = out$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Efull = out$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Efull = out$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point()

```

## Plot ISS V2.2
```{r}

df = tidytable(year = year,
               N10_Enull = out2$samples_10$expansion_NULL$length$iss$iss,
               N20_Enull = out2$samples_20$expansion_NULL$length$iss$iss,
               N30_Enull = out2$samples_30$expansion_NULL$length$iss$iss,
               N10_Eh = out2$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Eh = out2$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Eh = out2$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Ehm = out2$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Ehm = out2$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Ehm = out2$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point()

```