### 2024-12-10 Rmd/paper_plots/paper_plots_1.1.Rmd
paper runs and plots first commit

---
title: "Paper Plots V1"
author: "Brett Stacy"
date: "2024-11-25"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = "hide", fig.width = 8.5, fig.height = 5, dpi = 600)
library(tidytable)
library(data.table)
library(reshape2)
library(ggplot2)
library(knitr)
opts_knit$set(eval.after = "fig.cap")
options(scipen = 5)
```

# Introduction
Plots for paper:

1. Time series of new ISS vs old. Accompanied by number of hauls, nominal sample size, and possibly RSS boxplots. RSS boxplots may help justify focusing in on 20 sample size meidan for later plots to get an apples to apples, granular test of impact of sample size and expansion complexity.

2. ISS full, ISS without expansion, ISS without sampling strata, ISS without either.

3. 3x3 grid: sample size 10, 20, 30 fixed X expansion none, haul only, haul+month. With sampling strata.

3.supp.1: same as above but without sampling strata. 
3.supp.2: same as above but without bootstrapping trip. 
3.supp.3: same as above but without either.

## Plot 1.1
```{r}
out_1.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_1.1_ISS_100iters_output_2024-11-25.RDS")

out_2.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_2.1_ISS_100iters_output_2024-11-25.RDS")

out_3.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_3.1_ISS_100iters_output_2024-11-26.RDS")

out_4.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_4.1_ISS_100iters_output_2024-11-26.RDS")

out_5.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_5.1_ISS_100iters_output_2024-11-26.RDS")

out_6.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_6.1_ISS_100iters_output_2024-11-26.RDS")
out_7.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_7.1_ISS_100iters_output_2024-11-26.RDS")
out_8.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_8.1_ISS_100iters_output_2024-11-26.RDS")
out_9.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_9.1_ISS_100iters_output_2024-12-04.RDS")
out_10.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_10.1_ISS_100iters_output_2024-12-04.RDS")
out_11.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_11.1_ISS_100iters_output_2024-12-04.RDS")

assessment_2023_iss = tidytable(year = 1977:2023,
                                iss = c(27,
                                45,
                                54,
                                54,
                                63,
                                31,
                                117,
                                140,
                                211,
                                216,
                                396,
                                401,
                                185,
                                189,
                                1565,
                                1543,
                                913,
                                1435,
                                1561,
                                2020,
                                2142,
                                2046,
                                2722,
                                2960,
                                3140,
                                3446,
                                4301,
                                3608,
                                3455,
                                2694,
                                2115,
                                2502,
                                2218,
                                1975,
                                2596,
                                2776,
                                3293,
                                3604,
                                3324,
                                2821,
                                2469,
                                1899,
                                1367,
                                1048,
                                859,
                                1160,
                                706))
```

### baseISS with assessment ISS

```{r}
par(mfrow = c(1,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "ISS", xlab = "year", cex.lab = 1.3, ylim = c(0, max(out_1.1$length$iss$iss)*1.1), col = "blue") 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "black")
legend("topleft", lty = c(1,1), col = c("black", "blue"), legend = c("2023 assessment", "baseISS"))

```

### All together 4 panel

```{r, fig.height=8}
df = data.frame(rss = out_1.1$length$rss$rss, year = as.factor(out_1.1$length$rss$YEAR))

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
par(mfrow = c(4,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "ISS", xaxt = "n") 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue")
legend("topleft", lty = c(1,1), col = c("black", "blue"), legend = c("2023 assessment", "baseISS"))

plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$nhls, type = "l", ylab = "number of hauls", xlab = "", col = "black", cex.lab = 1.3, xaxt = "n")

plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$nss, type = "l", ylab = "nominal sample size", xlab = "", cex.lab = 1.3, xaxt = "n")

plot(df$year, df$rss, ylim = c(0, 20000), ylab = "RSS", cex.lab = 1.3, xlab = "year")

```

### All together 3 panel

```{r, fig.height=8}
df = data.frame(rss = out_1.1$length$rss$rss, year = as.factor(out_1.1$length$rss$YEAR))

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
par(mfrow = c(3,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "ISS", xaxt = "n", ylim = c(0, 15000)) 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue")
lines(out_1.1$length$iss$YEAR, out_1.1$length$iss$nhls, col = "red")
legend("topleft", lty = c(1,1,1), col = c("black", "blue", "red"), legend = c("2023 assessment", "baseISS", "number of hauls"))

plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$nss, type = "l", ylab = "nominal sample size", xlab = "", cex.lab = 1.3, xaxt = "n")

plot(df$year, df$rss, ylim = c(0, 20000), ylab = "RSS", cex.lab = 1.3, xlab = "year")

```

#### Inspect R-squared
Fit models to number of hauls.

```{r}
# baseISS ~ # hauls
fit1 = lm(out_1.1$length$iss$iss ~ out_1.1$length$iss$nhls)
summary(fit1)

# assessment ISS ~ # hauls
fit2 = lm(assessment_2023_iss$iss ~ out_1.1$length$iss$nhls)
summary(fit2)

# baseISS ~ assessment ISS
fit3 = lm(out_1.1$length$iss$iss ~ assessment_2023_iss$iss)
summary(fit3)

```

## Plot 2.1 
Deconstruct the contributors to baseISS

### All together 1

Subtracting things from baseISS

```{r, fig.height=8}

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
par(mfrow = c(5,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "baseISS", xaxt = "n") 
plot(out_2.1$length$iss$YEAR, out_2.1$length$iss$iss, type = "l", ylab = "baseISS - E", xaxt = "n") 
plot(out_3.1$length$iss$YEAR, out_3.1$length$iss$iss, type = "l", ylab = "baseISS - SS", xaxt = "n") 
plot(out_4.1$length$iss$YEAR, out_4.1$length$iss$iss, type = "l", ylab = "baseISS - (E,SS)", xaxt = "n") 
plot(out_5.1$length$iss$YEAR, out_5.1$length$iss$iss, type = "l", ylab = "baseISS - (E,SS,Bs)", cex.lab = 1.3, xlab = "year") 
```

### All together 2

Adding things to baseISS

```{r, fig.height=6}
par(mfrow = c(1,1))

plot(out_5.1$length$iss$YEAR, out_5.1$length$iss$iss, type = "l", ylab = "ISS", cex.lab = 1.3, col = "black")
lines(out_6.1$length$iss$YEAR, out_6.1$length$iss$iss, type = "l", col = "blue") 
lines(out_7.1$length$iss$YEAR, out_7.1$length$iss$iss, type = "l", col = "orange") 
lines(out_8.1$length$iss$YEAR, out_8.1$length$iss$iss, type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", col = "red")

legend("topleft", legend = c("-E,SS,Bs", "-SS,Bs", "-Bs", "-SS", "-0"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

#### sub range 1
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1999:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000))
lines(out_6.1$length$iss$YEAR[sub_range], out_6.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_7.1$length$iss$YEAR[sub_range], out_7.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_8.1$length$iss$YEAR[sub_range], out_8.1$length$iss$iss[sub_range], type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot length", "boot length and expand", "boot length, expand, sampling strata weighting", "boot length, expand, boot trip+haul", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

#### reorder long
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1977:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_9.1$length$iss$YEAR[sub_range], out_9.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_4.1$length$iss$YEAR[sub_range], out_4.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_2.1$length$iss$YEAR[sub_range], out_2.1$length$iss$iss[sub_range], type = "l", col = "green")
# lines(out_8.1$length$iss$YEAR[sub_range], out_8.1$length$iss$iss[sub_range], type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot length", "boot length, haul", "boot all", "boot all, sampling strata weighting", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

#### reorder short
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1998:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_9.1$length$iss$YEAR[sub_range], out_9.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_4.1$length$iss$YEAR[sub_range], out_4.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_2.1$length$iss$YEAR[sub_range], out_2.1$length$iss$iss[sub_range], type = "l", col = "green")
# lines(out_8.1$length$iss$YEAR[sub_range], out_8.1$length$iss$iss[sub_range], type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot length", "boot length, haul", "boot all", "boot all, sampling strata weighting", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

#### reorder short 2
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1998:2023

par(mfrow = c(1,1))
plot(out_10.1$length$iss$YEAR[sub_range], out_10.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_11.1$length$iss$YEAR[sub_range], out_11.1$length$iss$iss[sub_range], type = "l", col = "blue")

lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot.length", "boot.length + boot.haul", "boot.length + boot.haul + boot.trip"), lty = c(1,1,1), col = c("black", "blue", "red"))
```

## Plot 3.1

3x3 grid.

### Data
```{r}
out_2.1.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_2.1.1_ISS_200_iters_2024-11-27.RDS")

out_2.2.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_2.2.1_ISS_200_iters_2024-12-02.RDS")

year = out_2.1.1$samples_10$expansion_NULL$length$iss$YEAR

samples_global = c(10, 20, 30)
expansion_global = list(NULL, "haul_numbers", c("haul_numbers", "month_numbers"))
expansion_names_global = c("NULL", "haul_numbers", c("haul_month_numbers"))

```

### With sampling strata
```{r}

df = tidytable(year = year,
               N10_Enone = out_2.1.1$samples_10$expansion_NULL$length$iss$iss,
               N20_Enone = out_2.1.1$samples_20$expansion_NULL$length$iss$iss,
               N30_Enone = out_2.1.1$samples_30$expansion_NULL$length$iss$iss,
               N10_Emedium = out_2.1.1$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out_2.1.1$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Emedium = out_2.1.1$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Efull = out_2.1.1$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out_2.1.1$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Efull = out_2.1.1$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point()

```

### Without sampling strata
```{r}

df = tidytable(year = year,
               N10_Enone = out_2.2.1$samples_10$expansion_NULL$length$iss$iss,
               N20_Enone = out_2.2.1$samples_20$expansion_NULL$length$iss$iss,
               N30_Enone = out_2.2.1$samples_30$expansion_NULL$length$iss$iss,
               N10_Emedium = out_2.2.1$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out_2.2.1$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Emedium = out_2.2.1$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Efull = out_2.2.1$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out_2.2.1$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Efull = out_2.2.1$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point()

```