### 2024-12-18 Rmd/paper_plots/paper_plots_1.1.Rmd
run Trip and Haul number variation scenarios and plot.

---
title: "Paper Plots V1"
author: "Brett Stacy"
date: "2024-11-25"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = "hide", fig.width = 8.5, fig.height = 5, dpi = 600)
library(tidytable)
library(data.table)
library(reshape2)
library(ggplot2)
library(knitr)
opts_knit$set(eval.after = "fig.cap")
options(scipen = 5)
```

# Introduction
Plots for paper:

1. Time series of new ISS vs old. Accompanied by number of hauls, nominal sample size, and possibly RSS boxplots. RSS boxplots may help justify focusing in on 20 sample size meidan for later plots to get an apples to apples, granular test of impact of sample size and expansion complexity.

2. ISS full, ISS without expansion, ISS without sampling strata, ISS without either.

3. 3x3 grid: sample size 10, 20, 30 fixed X expansion none, haul only, haul+month. With sampling strata.

3.supp.1: same as above but without sampling strata. 
3.supp.2: same as above but without bootstrapping trip. 
3.supp.3: same as above but without either.

## Plot 1.1
```{r}
out_1.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_1.1_ISS_100iters_output_2024-11-25.RDS")

out_2.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_2.1_ISS_100iters_output_2024-11-25.RDS")

out_3.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_3.1_ISS_100iters_output_2024-11-26.RDS")

out_4.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_4.1_ISS_100iters_output_2024-11-26.RDS")

out_5.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_5.1_ISS_100iters_output_2024-11-26.RDS")

out_6.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_6.1_ISS_100iters_output_2024-11-26.RDS")
out_7.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_7.1_ISS_100iters_output_2024-11-26.RDS")
out_8.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_8.1_ISS_100iters_output_2024-11-26.RDS")
out_9.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_9.1_ISS_100iters_output_2024-12-04.RDS")
out_10.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_10.1_ISS_100iters_output_2024-12-04.RDS")
out_11.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_11.1_ISS_100iters_output_2024-12-04.RDS")
out_12.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_12.1_ISS_100iters_output_2024-12-10.RDS")
out_13.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_13.1_ISS_100iters_output_2024-12-10.RDS")
out_14.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_14.1_ISS_100iters_output_2024-12-10.RDS")
out_15.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_15.1_ISS_100iters_output_2024-12-10.RDS")
out_16.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_16.1_ISS_100iters_output_2024-12-11.RDS")
out_17.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_17.1_ISS_100iters_output_2024-12-11.RDS")

assessment_2023_iss = tidytable(year = 1977:2023,
                                iss = c(27,
                                45,
                                54,
                                54,
                                63,
                                31,
                                117,
                                140,
                                211,
                                216,
                                396,
                                401,
                                185,
                                189,
                                1565,
                                1543,
                                913,
                                1435,
                                1561,
                                2020,
                                2142,
                                2046,
                                2722,
                                2960,
                                3140,
                                3446,
                                4301,
                                3608,
                                3455,
                                2694,
                                2115,
                                2502,
                                2218,
                                1975,
                                2596,
                                2776,
                                3293,
                                3604,
                                3324,
                                2821,
                                2469,
                                1899,
                                1367,
                                1048,
                                859,
                                1160,
                                706))
```

### baseISS with assessment ISS

```{r}
par(mfrow = c(1,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "ISS", xlab = "year", cex.lab = 1.3, ylim = c(0, max(out_1.1$length$iss$iss)*1.1), col = "blue") 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "black")
legend("topleft", lty = c(1,1), col = c("black", "blue"), legend = c("2023 assessment", "baseISS"))

```

### All together 4 panel

```{r, fig.height=8}
df = data.frame(rss = out_1.1$length$rss$rss, year = as.factor(out_1.1$length$rss$YEAR))

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
par(mfrow = c(4,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "ISS", xaxt = "n") 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue")
legend("topleft", lty = c(1,1), col = c("black", "blue"), legend = c("2023 assessment", "baseISS"))

plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$nhls, type = "l", ylab = "number of hauls", xlab = "", col = "black", cex.lab = 1.3, xaxt = "n")

plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$nss, type = "l", ylab = "nominal sample size", xlab = "", cex.lab = 1.3, xaxt = "n")

plot(df$year, df$rss, ylim = c(0, 20000), ylab = "RSS", cex.lab = 1.3, xlab = "year")

```

### All together 3 panel

```{r, fig.height=8}
df = data.frame(rss = out_1.1$length$rss$rss, year = as.factor(out_1.1$length$rss$YEAR))

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
# par(mfrow = c(3,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
par(mfrow = c(3,1), mar = c(0,3.1,0,0), oma = c(2, 0, 2, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "ISS", xaxt = "n", ylim = c(0, 15000)) 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue")
lines(out_1.1$length$iss$YEAR, out_1.1$length$iss$nhls, col = "red")
legend("topleft", lty = c(1,1,1), col = c("black", "blue", "red"), legend = c("ISSB", "ISSA", "number of hauls"))

plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$nss, type = "l", ylab = "nominal sample size", xlab = "", cex.lab = 1.3, xaxt = "n")

plot(df$year, df$rss, ylim = c(0, 20000), ylab = "RSS", cex.lab = 1.3)

```

#### Inspect R-squared
Fit models to number of hauls.

```{r}
# baseISS ~ # hauls
fit1 = lm(out_1.1$length$iss$iss ~ out_1.1$length$iss$nhls)
summary(fit1)

# assessment ISS ~ # hauls
fit2 = lm(assessment_2023_iss$iss ~ out_1.1$length$iss$nhls)
summary(fit2)

# baseISS ~ assessment ISS
fit3 = lm(out_1.1$length$iss$iss ~ assessment_2023_iss$iss)
summary(fit3)

```

## Plot 2.1 
Deconstruct the contributors to baseISS

### All together 1

Subtracting things from baseISS

```{r, fig.height=8}

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
par(mfrow = c(5,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
plot(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", ylab = "baseISS", xaxt = "n") 
plot(out_2.1$length$iss$YEAR, out_2.1$length$iss$iss, type = "l", ylab = "baseISS - E", xaxt = "n") 
plot(out_3.1$length$iss$YEAR, out_3.1$length$iss$iss, type = "l", ylab = "baseISS - SS", xaxt = "n") 
plot(out_4.1$length$iss$YEAR, out_4.1$length$iss$iss, type = "l", ylab = "baseISS - (E,SS)", xaxt = "n") 
plot(out_5.1$length$iss$YEAR, out_5.1$length$iss$iss, type = "l", ylab = "baseISS - (E,SS,Bs)", cex.lab = 1.3, xlab = "year") 
```

### All together 2

Adding things to baseISS

```{r, fig.height=6}
par(mfrow = c(1,1))

plot(out_5.1$length$iss$YEAR, out_5.1$length$iss$iss, type = "l", ylab = "ISS", cex.lab = 1.3, col = "black")
lines(out_6.1$length$iss$YEAR, out_6.1$length$iss$iss, type = "l", col = "blue") 
lines(out_7.1$length$iss$YEAR, out_7.1$length$iss$iss, type = "l", col = "orange") 
lines(out_8.1$length$iss$YEAR, out_8.1$length$iss$iss, type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR, out_1.1$length$iss$iss, type = "l", col = "red")

legend("topleft", legend = c("-E,SS,Bs", "-SS,Bs", "-Bs", "-SS", "-0"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

#### sub range 1
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1999:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000))
lines(out_6.1$length$iss$YEAR[sub_range], out_6.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_7.1$length$iss$YEAR[sub_range], out_7.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_8.1$length$iss$YEAR[sub_range], out_8.1$length$iss$iss[sub_range], type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot length", "boot length and expand", "boot length, expand, sampling strata weighting", "boot length, expand, boot trip+haul", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

#### reorder long

\newpage

```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1977:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISSB", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_9.1$length$iss$YEAR[sub_range], out_9.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_4.1$length$iss$YEAR[sub_range], out_4.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_2.1$length$iss$YEAR[sub_range], out_2.1$length$iss$iss[sub_range], type = "l", col = "green")
# lines(out_8.1$length$iss$YEAR[sub_range], out_8.1$length$iss$iss[sub_range], type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topleft", legend = c("boot length", "boot length + boot haul", "boot all", "boot all + sampling strata weighting", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

\newpage

#### reorder short

\newpage

```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1999:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISSB", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "")
lines(out_9.1$length$iss$YEAR[sub_range], out_9.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_4.1$length$iss$YEAR[sub_range], out_4.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_2.1$length$iss$YEAR[sub_range], out_2.1$length$iss$iss[sub_range], type = "l", col = "green")
# lines(out_8.1$length$iss$YEAR[sub_range], out_8.1$length$iss$iss[sub_range], type = "l", col = "green") 
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot length", "boot length + boot haul", "boot all", "boot all + sampling strata weighting", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"))
```

\newpage

#### reorder short with half trips and half hauls
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1998:2023

par(mfrow = c(1,1))
plot(out_5.1$length$iss$YEAR[sub_range], out_5.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_9.1$length$iss$YEAR[sub_range], out_9.1$length$iss$iss[sub_range], type = "l", col = "blue") 
lines(out_4.1$length$iss$YEAR[sub_range], out_4.1$length$iss$iss[sub_range], type = "l", col = "orange") 
lines(out_12.1$length$iss$YEAR[sub_range], out_12.1$length$iss$iss[sub_range], type = "l", col = "purple")
lines(out_13.1$length$iss$YEAR[sub_range], out_13.1$length$iss$iss[sub_range], type = "l", col = "brown")
lines(out_2.1$length$iss$YEAR[sub_range], out_2.1$length$iss$iss[sub_range], type = "l", col = "green")
lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot length", "boot length, haul", "boot all", "boot all, half trips", "boot all, half hauls", "boot all, sampling strata weighting", "full"), lty = c(1,1,1,1,1,1,1), col = c("black", "blue", "orange","purple", "brown", "green", "red"))
```

#### reorder short 2
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1998:2023

par(mfrow = c(1,1))
plot(out_10.1$length$iss$YEAR[sub_range], out_10.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_11.1$length$iss$YEAR[sub_range], out_11.1$length$iss$iss[sub_range], type = "l", col = "blue")

lines(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("boot.length", "boot.length + boot.haul", "boot.length + boot.haul + boot.trip"), lty = c(1,1,1), col = c("black", "blue", "red"))
```

#### Full, half trip, half haul
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1998:2023

par(mfrow = c(1,1))
plot(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_14.1$length$iss$YEAR[sub_range], out_14.1$length$iss$iss[sub_range], type = "l", col = "blue")

lines(out_15.1$length$iss$YEAR[sub_range], out_15.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("Full", "Half Trip", "Half Haul"), lty = c(1,1,1), col = c("black", "blue", "red"))
```

#### Full, 1.5 trip, 1.5 haul
```{r, fig.height=6}
sub_range = out_5.1$length$iss$YEAR %in% 1998:2023

par(mfrow = c(1,1))
plot(out_1.1$length$iss$YEAR[sub_range], out_1.1$length$iss$iss[sub_range], type = "l", ylab = "ISS", cex.lab = 1.3, col = "black", ylim = c(0, 25000), xlab = "Year")
lines(out_16.1$length$iss$YEAR[sub_range], out_16.1$length$iss$iss[sub_range], type = "l", col = "blue")

lines(out_17.1$length$iss$YEAR[sub_range], out_17.1$length$iss$iss[sub_range], type = "l", col = "red")

legend("topright", legend = c("Full", "1.5 Trip", "1.5 Haul"), lty = c(1,1,1), col = c("black", "blue", "red"))
```

## Plot 7x2 Grid Haul and Trip numbers

### Trip Change
```{r}

out_18.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_18.1_ISS_200_iters_2024-12-12.RDS")

year = out_18.1$proportion_0.25$type_trip$length$iss$YEAR
proportion_global = c(0.25, 0.50, 0.75, 1.25, 1.50, 1.75, 2.00)
type_global = list("trip", "haul")
type_names_global = c("trip", "haul")

df = tidytable(year = year,
               N_0.25 = out_18.1$proportion_0.25$type_trip$length$iss$iss,
               N_0.50 = out_18.1$proportion_0.5$type_trip$length$iss$iss,
               N_0.75 = out_18.1$proportion_0.75$type_trip$length$iss$iss,
               N_1.00 = out_1.1$length$iss$iss,
               N_1.25 = out_18.1$proportion_1.25$type_trip$length$iss$iss,
               N_1.50 = out_18.1$proportion_1.5$type_trip$length$iss$iss,
               N_1.75 = out_18.1$proportion_1.75$type_trip$length$iss$iss,
               N_2.00 = out_18.1$proportion_2$type_trip$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point() +
  ylim(0, 17000) +
  xlim(1998, 2023)

```

### Boxplots
REMEMBER! these are the full year range, may want to narrow it to recent range.
```{r}
df2 %>% 
  tidytable::mutate(ISS.diff = (df2$ISS[df2$scenario=="N_1.00"]) - ISS,
                    ISS.error = (ISS - df2$ISS[df2$scenario=="N_1.00"])/df2$ISS[df2$scenario=="N_1.00"],
                    type = "trip",
                    .by = scenario) -> df3.1

# plot(df3$scenario, df3$ISS.diff)
plot(df3.1$scenario, df3.1$ISS.error)
abline(h = c(-.1,.1), lty = "dashed")

```

### Haul Change
```{r}

out_18.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_18.1_ISS_200_iters_2024-12-12.RDS")

year = out_18.1$proportion_0.25$type_trip$length$iss$YEAR
proportion_global = c(0.25, 0.50, 0.75, 1.25, 1.50, 1.75, 2.00)
type_global = list("trip", "haul")
type_names_global = c("trip", "haul")

df = tidytable(year = year,
               N_0.25 = out_18.1$proportion_0.25$type_haul$length$iss$iss,
               N_0.50 = out_18.1$proportion_0.5$type_haul$length$iss$iss,
               N_0.75 = out_18.1$proportion_0.75$type_haul$length$iss$iss,
               N_1.00 = out_1.1$length$iss$iss,
               N_1.25 = out_18.1$proportion_1.25$type_haul$length$iss$iss,
               N_1.50 = out_18.1$proportion_1.5$type_haul$length$iss$iss,
               N_1.75 = out_18.1$proportion_1.75$type_haul$length$iss$iss,
               N_2.00 = out_18.1$proportion_2$type_haul$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point() +
  ylim(0, 17000) +
  xlim(1998, 2023)

```

### Boxplots
```{r}
df2 %>% 
  tidytable::mutate(ISS.diff = (df2$ISS[df2$scenario=="N_1.00"]) - ISS,
                    ISS.error = (ISS - df2$ISS[df2$scenario=="N_1.00"])/df2$ISS[df2$scenario=="N_1.00"],
                    type = "haul",
                    .by = scenario) -> df3.2

# plot(df3$scenario, df3$ISS.diff)
plot(df3.2$scenario, df3.2$ISS.error)
abline(h = c(-.1,.1), lty = "dashed")

```

### Boxplots Together 1977-2023
```{r, fig.width=8}

df3.1 %>% 
  tidytable::bind_rows(df3.2) -> df4

df4 %>% 
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Scenario", y = "ISSB") +
  theme_classic() +
  theme(legend.position = c(0.5, 0.9))

```

### Boxplots Together 1999-2023
```{r}

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>% 
  tidytable::filter(year %in% 1999:2023) -> df4

df4 %>% 
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type))

```

## Plot 3.1

3x3 grid.

### Data
```{r}
out_2.1.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_2.1.1_ISS_200_iters_2024-11-27.RDS")

out_2.2.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_2.2.1_ISS_200_iters_2024-12-02.RDS")

year = out_2.1.1$samples_10$expansion_NULL$length$iss$YEAR

samples_global = c(10, 20, 30)
expansion_global = list(NULL, "haul_numbers", c("haul_numbers", "month_numbers"))
expansion_names_global = c("NULL", "haul_numbers", c("haul_month_numbers"))

```

### With sampling strata
```{r, fig.width=9, fig.height=5}
df = tidytable(year = year,
               N10_Enone = out_2.1.1$samples_10$expansion_NULL$length$iss$iss,
               N20_Enone = out_2.1.1$samples_20$expansion_NULL$length$iss$iss,
               N30_Enone = out_2.1.1$samples_30$expansion_NULL$length$iss$iss,
               N10_Emedium = out_2.1.1$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out_2.1.1$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Emedium = out_2.1.1$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Efull = out_2.1.1$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out_2.1.1$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Efull = out_2.1.1$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "", y = "ISSB") +
  geom_point() + 
  theme_classic() +
  theme(legend.position = c(0.84, 0.8),
        axis.title.x = element_blank())
```

```{r, fig.width=9, fig.height=5}
# stats
((df$N10_Enone - df$N20_Enone)/df$N20_Enone) %>%
  mean()

((df$N10_Emedium - df$N20_Emedium)/df$N20_Emedium) %>%
  mean()

((df$N10_Efull - df$N20_Efull)/df$N20_Efull) %>%
  mean()
(-.02 - .12 - .08)/3

((df$N30_Enone - df$N20_Enone)/df$N20_Enone) %>%
  mean()

((df$N30_Emedium - df$N20_Emedium)/df$N20_Emedium) %>%
  mean()

((df$N30_Efull - df$N20_Efull)/df$N20_Efull) %>%
  mean()

(.03 + .04 +.03)/3

```

### Without sampling strata
```{r}

df = tidytable(year = year,
               N10_Enone = out_2.2.1$samples_10$expansion_NULL$length$iss$iss,
               N20_Enone = out_2.2.1$samples_20$expansion_NULL$length$iss$iss,
               N30_Enone = out_2.2.1$samples_30$expansion_NULL$length$iss$iss,
               N10_Emedium = out_2.2.1$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out_2.2.1$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Emedium = out_2.2.1$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Efull = out_2.2.1$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out_2.2.1$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Efull = out_2.2.1$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "Year", y = "ISS") +
  geom_point()

```

## Plot 4.1 Trip and Haul x .5, 1.5

2x3 grid.

### Data
```{r}
out_19.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_19.1_ISS_200_iters_2024-12-17.RDS")

out_20.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_20.1_ISS_200_iters_2024-12-17.RDS")

year = out_19.1$proportion_0.5$expansion_NULL$length$iss$YEAR

proportion_global = c(0.50, 1.50)
expansion_global = list(NULL, "haul_numbers", c("haul_numbers", "month_numbers"))
expansion_names_global = c("NULL", "haul_numbers", c("haul_month_numbers"))

samples_global = c(10, 20, 30)
expansion_global = list(NULL, "haul_numbers", c("haul_numbers", "month_numbers"))
expansion_names_global = c("NULL", "haul_numbers", c("haul_month_numbers"))

```

COME BACK TO THIS! NEED TO RERUN WITH REGULAR HAUL AND TRIP NUMBERS!
### With sampling strata
```{r, fig.width=9, fig.height=5}
df = tidytable(year = year,
               N0.5_Enone = out_19.1$proportion_0.5$expansion_NULL$length$iss$iss,
               N1.0_Enone = out_2.1.1$samples_20$expansion_NULL$length$iss$iss,
               N30_Enone = out_2.1.1$samples_30$expansion_NULL$length$iss$iss,
               N10_Emedium = out_2.1.1$samples_10$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out_2.1.1$samples_20$expansion_haul_numbers$length$iss$iss,
               N30_Emedium = out_2.1.1$samples_30$expansion_haul_numbers$length$iss$iss,
               N10_Efull = out_2.1.1$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out_2.1.1$samples_20$expansion_haul_month_numbers$length$iss$iss,
               N30_Efull = out_2.1.1$samples_30$expansion_haul_month_numbers$length$iss$iss,)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  labs(x = "", y = "ISSB") +
  geom_point() + 
  theme_classic() +
  theme(legend.position = c(0.84, 0.8),
        axis.title.x = element_blank())
```