### 2025-03-31 Rmd/paper_plots/paper_plots_1.5.Rmd

---
title: "Paper Plots V5"
author: "Brett Stacy"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, warning = F, message = F, results = "hide", fig.width = 8.5, fig.height = 5, dpi = 600)
library(tidytable)
library(data.table)
library(reshape2)
library(ggplot2)
library(knitr)
library(stringr)
library(patchwork)
library(viridis)
library(ggpattern)
opts_knit$set(eval.after = "fig.cap")
options(scipen = 5)
```

# Introduction
Plots for paper:

## Data
```{r}

out_1.4.old = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_1.4_ISS_1000iters_output_2025-01-23.RDS")

out_1.5 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_1.5_ISS_1000iters_output_2025-03-24.RDS")

# out_3.2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_3.2_ISS_500iters_output_2025-03-07.RDS") ### mislabeled

out_5.2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_41.1_ISS_500iters_output_2025-03-24.RDS")

out_9.2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_42.1_ISS_500iters_output_2025-03-24.RDS")

out_4.2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_43.1_ISS_500iters_output_2025-03-24.RDS")
# 
# out_31.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_31.1.TAS_ISS_500_iters_2025-03-02.RDS")
# 
# out_39.5 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_39.5_ISS_500_iters_2025-03-21.RDS")
#

out_37.2 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_37.2_ISS_500iters_output_2025-03-26.RDS")
out_37.3 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_37.3_ISS_500iters_output_2025-03-26.RDS")

# save out_1.5 for Steve to plug into the assessment
out_1.5$length$iss %>% 
  select(YEAR, iss) %>% 
  write.csv(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/Fishery ISS Paper/issb_v1_from_Brett_Stacy.csv", row.names = F)

assessment_2023_iss = tidytable(year = 1977:2023,
                                iss = c(27,
                                45,
                                54,
                                54,
                                63,
                                31,
                                117,
                                140,
                                211,
                                216,
                                396,
                                401,
                                185,
                                189,
                                1565,
                                1543,
                                913,
                                1435,
                                1561,
                                2020,
                                2142,
                                2046,
                                2722,
                                2960,
                                3140,
                                3446,
                                4301,
                                3608,
                                3455,
                                2694,
                                2115,
                                2502,
                                2218,
                                1975,
                                2596,
                                2776,
                                3293,
                                3604,
                                3324,
                                2821,
                                2469,
                                1899,
                                1367,
                                1048,
                                859,
                                1160,
                                706))

# survey vs observer length measurements
obs_numbers = tidytable(year = 1982:2023,
                        survey = c(10863,
                        13143,
                        12133,
                        17150,
                        15872,
                        9483,
                        6950,
                        4246,
                        5428,
                        7069,
                        10129,
                        10500,
                        12931,
                        9820,
                        9348,
                        9591,
                        9574,
                        11183,
                        12170,
                        19078,
                        12365,
                        11835,
                        10986,
                        11753,
                        12530,
                        13441,
                        15328,
                        23737,
                        21223,
                        25150,
                        30177,
                        19902,
                        29204,
                        19880,
                        19507,
                        15020,
                        8806,
                        23408,
                        0,
                        17397,
                        16677,
                        19943),
                        fishery = c(9658,
                                    33200,
                                    45635,
                                    66940,
                                    58257,
                                    129226,
                                    111065,
                                    58625,
                                    39698,
                                    347227,
                                    344923,
                                    248967,
                                    359147,
                                    344794,
                                    445217,
                                    474908,
                                    438746,
                                    186233,
                                    199708,
                                    210419,
                                    230802,
                                    288854,
                                    237487,
                                    228664,
                                    179782,
                                    140663,
                                    164860,
                                    147875,
                                    131514,
                                    172269,
                                    192273,
                                    211962,
                                    234476,
                                    213888,
                                    182980,
                                    157482,
                                    124004,
                                    86800,
                                    65301,
                                    55858,
                                    73025,
                                    42942))

obs_numbers %>% 
  tidytable::summarise(sum(survey), sum(fishery))

obs_numbers %>% 
  tidytable::summarise(mean(survey), mean(fishery))

```

## ISSB Results 
### * Time series 2 panel

```{r, fig.height=6}
df = data.frame(rss = out_1.5$length$rss$rss, year = as.factor(out_1.5$length$rss$YEAR))

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
# par(mfrow = c(3,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
par(mfrow = c(2,1), mar = c(0,3.1,0,0), oma = c(3, 0, 2, .5), mgp = c(2, .6, 0))
plot(out_1.5$length$iss$YEAR, out_1.5$length$iss$iss, type = "o", pch = 19, ylab = "ISS", xaxt = "n", ylim = c(0, 15000), cex.lab = 1.3) 
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue", type = "o", pch = 3)
lines(out_1.5$length$iss$YEAR, out_1.5$length$iss$nhls, col = "orange", type = "o", pch = 5)
legend(1977, 11000, lty = c(1,1,1), col = c("black", "blue", "orange"), pch = c(19, 3, 5), legend = c("ISSB", "ISSA", "number of hauls"), bty = "n")
text(1977, 14000, "A")

plot(out_1.5$length$iss$YEAR, out_1.5$length$iss$nss, type = "l", ylab = "Nominal Sample Size", xlab = "Year", cex.lab = 1.3) 
text(1977, 420000, "B")
mtext("Year", side = 1, line = 2, cex = 1.5)

# compare with and without SS
par(mfrow = c(1,1))
plot(out_1.5$length$iss$YEAR, out_1.5$length$iss$iss, type = "o", pch = 19, ylab = "ISSB", ylim = c(0, 15000), cex.lab = 1.3) 
lines(out_1.4.old$length$iss$YEAR, out_1.4.old$length$iss$iss, col = "blue", type = "o", pch = 5)
legend(1977, 11000, lty = c(1,1), col = c("black", "blue"), pch = c(19, 5), legend = c("Not including sampling strata weighting", "Including sampling strata weighting"), bty = "n")

# averages
out_1.5$length$iss$iss %>% mean()
assessment_2023_iss$iss %>% mean()

```

### * Time seris components

\newpage

```{r, fig.height=5.5}
sub_range = out_5.2$length$iss$YEAR %in% 1977:2023

par(mfrow = c(1,1))
plot(out_5.2$length$iss$YEAR[sub_range], out_5.2$length$iss$iss[sub_range], ylab = "ISSB", cex.lab = 1.3, col = "cyan", ylim = c(0, 25000), xlab = "Year", lwd = 2, type = "o", pch= 18)
lines(out_9.2$length$iss$YEAR[sub_range], out_9.2$length$iss$iss[sub_range], type = "o", pch = 15, col = "orange3", lwd = 2)
lines(out_4.2$length$iss$YEAR[sub_range], out_4.2$length$iss$iss[sub_range], type = "o", pch = 17, col = "magenta", lwd = 2)
lines(out_1.5$length$iss$YEAR[sub_range], out_1.5$length$iss$iss[sub_range], type = "o", pch = 19, col = "black", lwd = 2, lty = 1)

# legend("topleft", title = "Component Case", legend = c("1", "2", "3", "4", "full"), lty = c(1,1,1,1,1), col = c("black", "blue", "orange", "green", "red"), bty = "n")
legend("topleft", title = "Component Case", legend = c("1", "2", "3", "full"), lty = c(1,1,1,1), pch = c(18, 15, 17, 19), col = c("cyan", "orange3", "magenta", "black"), bty = "n", lwd = rep(2, 4))

# not accounting for correlations exceed hauls?
sum((out_5.2$length$iss$iss[sub_range]-out_1.5$length$iss$nhls) < 0)

```

\newpage

#### * TOGETHER

\newpage

```{r, fig.height=8}
df = data.frame(rss = out_1.5$length$rss$rss, year = as.factor(out_1.5$length$rss$YEAR))

# par(mfrow = c(4,1), mar = c(0,0,0,0), oma = c(2, 4, 2, .5), mgp = c(2, .6, 0))
# par(mfrow = c(3,1), mar = c(1.5, 3.1, 0, .5), mgp = c(2, .6, 0))
par(mfrow = c(3,1), mar = c(0,3.1,0,0), oma = c(3, 0, 2, .5), mgp = c(2, .6, 0))
plot(out_1.5$length$iss$YEAR, out_1.5$length$iss$iss, type = "o", pch = 19, ylab = "ISS", xaxt = "n", ylim = c(0, 15000), cex.lab = 1.5, cex.axis = 1.3)
lines(assessment_2023_iss$year, assessment_2023_iss$iss, col = "blue", type = "o", pch = 3)
lines(out_1.5$length$iss$YEAR, out_1.5$length$iss$nhls, col = "orange", type = "o", 5)
legend(1977, 11000, lty = c(1,1,1), col = c("black", "blue", "orange"), pch = c(19, 3, 5), legend = c("ISSB", "ISSA", "number of hauls"), bty = "n", cex = 1.5)
text(1977, 14000, "A", cex = 1.5)

plot(out_1.5$length$iss$YEAR, out_1.5$length$iss$nss, type = "l", ylab = "Nominal Sample Size", xaxt = "n", cex.lab = 1.5, cex.axis = 1.3)
text(1977, 420000, "B", cex = 1.5)

plot(out_5.2$length$iss$YEAR[sub_range], out_5.2$length$iss$iss[sub_range], ylab = "ISSB", cex.lab = 1.5, col = "cyan", ylim = c(0, 27000), xlab = "", xaxt = "n", lwd = 1.5, type = "o", pch= 18, cex.axis = 1.3)
axis(side = 1, at = seq(1980, 2020, by = 5), cex.axis = 1.3)
lines(out_9.2$length$iss$YEAR[sub_range], out_9.2$length$iss$iss[sub_range], type = "o", pch = 15, col = "orange3", lwd = 1.5)
lines(out_4.2$length$iss$YEAR[sub_range], out_4.2$length$iss$iss[sub_range], type = "o", pch = 17, col = "magenta", lwd = 1.5)
lines(out_1.5$length$iss$YEAR[sub_range], out_1.5$length$iss$iss[sub_range], type = "o", pch = 19, col = "black", lwd = 1.5, lty = 1)
legend(1977, 20000, title = "Scenario", legend = c("1", "2", "3", "Full"), lty = c(1,1,1,1), pch = c(18, 15, 17, 19), col = c("cyan", "orange3", "magenta", "black"), bty = "n", lwd = rep(1.5, 4), cex = 1.5)
text(1977, 23000, "C", cex = 1.5)

# mtext("Year", side = 1, line = 2, cex = 1.3)

```

\newpage

# * 2003 & 2004 Histograms

\newpage

```{r, fig.height=10}
out_1.5$length$og_props %>% 
  tidytable::filter(YEAR %in% 2003:2004) %>% 
  tidytable::mutate(Year = as.factor(YEAR)) -> props.sub

p1 = ggplot(data = props.sub, aes(x = LENGTH, y = og_FREQ, colour = Year)) +
  geom_line() +
  geom_area(aes(fill = Year, group = Year), alpha = 0.5, position = "identity") +
  # geom_area_pattern(aes(pattern = Year, y = og_FREQ, x = LENGTH), alpha = 0.5, position = "identity") +
  theme_classic() +
  scale_x_continuous(limits = c(22, 108),
                     n.breaks = 9) +
  labs(x = "Length (cm)", y = "Frequency", tag = "A") +
  theme(legend.position = c(0.8, 0.8),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) +
  scale_y_continuous(limits = c(0, .04), expand = c(0, 0)) +
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# now do RSS
out_1.5$length$rss %>% 
  tidytable::filter(YEAR %in% 2003:2004) %>% 
  tidytable::mutate(Year = as.factor(YEAR)) -> props.sub

p3 = ggplot(data = props.sub, aes(x = rss, colour = Year)) +
  geom_histogram(aes(fill = Year, group = Year), alpha = 0.5, position = "identity", binwidth = 200) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) +
  labs(x = "RSS", y = "Count", tag = "E") + 
  geom_point(data = props.sub %>% 
               tidytable::left_join(out_1.5$length$iss %>% 
                                      tidytable::select(YEAR, iss)) %>% 
               tidytable::summarise(x = unique(iss), .by = Year) %>% 
               # tidytable::summarise(x = median(rss), .by = Year) %>% 
               tidytable::mutate(y = c(6,6)), aes(x,y, fill = Year), shape = 25, size = 4, color = "black") +
  scale_y_continuous(limits = c(0, 75), expand = c(0, 0)) +
  # scale_x_continuous(n.breaks = 8)
  scale_x_continuous(breaks = seq(0, 35000, by = 2500),
                     labels = c(0, "", 5000, "", 10000, "", 15000, "", 20000, "", 25000, "", 30000, "", 35000),
                     limits = c(0, 35000),
                     expand = c(0, 0))+
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# now do RSS with boot_length only.
out_5.2$length$rss %>% 
  tidytable::filter(YEAR %in% 2003:2004) %>% 
  tidytable::mutate(Year = as.factor(YEAR)) -> props.sub

p3.2 = ggplot(data = props.sub, aes(x = rss, colour = Year)) +
  geom_histogram(aes(fill = Year, group = Year), alpha = 0.5, position = "identity", binwidth = 200) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) +
  labs(x = "RSS", y = "Count", tag = "B") + 
  geom_point(data = props.sub %>% 
               tidytable::left_join(out_5.2$length$iss %>% 
                                      tidytable::select(YEAR, iss)) %>% 
               tidytable::summarise(x = unique(iss), .by = Year) %>% 
               # tidytable::summarise(x = median(rss), .by = Year) %>% 
               tidytable::mutate(y = c(9,9)), aes(x,y, fill = Year), shape = 25, size = 4, color = "black") +
  scale_y_continuous(limits = c(0, 110), expand = c(0, 0)) +
  # scale_x_continuous(n.breaks = 8)
  scale_x_continuous(breaks = seq(0, 35000, by = 2500),
                     labels = c(0, "", 5000, "", 10000, "", 15000, "", 20000, "", 25000, "", 30000, "", 35000),
                     limits = c(0, 35000),
                     expand = c(0, 0))+
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# now do RSS with boot_length and boot_haul.
out_9.2$length$rss %>% 
  tidytable::filter(YEAR %in% 2003:2004) %>% 
  tidytable::mutate(Year = as.factor(YEAR)) -> props.sub

p3.3 = ggplot(data = props.sub, aes(x = rss, colour = Year)) +
  geom_histogram(aes(fill = Year, group = Year), alpha = 0.5, position = "identity", binwidth = 200) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) +
  labs(x = "RSS", y = "Count", tag = "C") + 
  geom_point(data = props.sub %>% 
               tidytable::left_join(out_9.2$length$iss %>% 
                                      tidytable::select(YEAR, iss)) %>% 
               tidytable::summarise(x = unique(iss), .by = Year) %>% 
               # tidytable::summarise(x = median(rss), .by = Year) %>% 
               tidytable::mutate(y = c(6,6)), aes(x,y, fill = Year), shape = 25, size = 4, color = "black") +
  scale_y_continuous(limits = c(0, 75), expand = c(0, 0)) +
  # scale_x_continuous(n.breaks = 8)
  scale_x_continuous(breaks = seq(0, 35000, by = 2500),
                     labels = c(0, "", 5000, "", 10000, "", 15000, "", 20000, "", 25000, "", 30000, "", 35000),
                     limits = c(0, 35000),
                     expand = c(0, 0))+
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# now do RSS with boot all.
out_4.2$length$rss %>% 
  tidytable::filter(YEAR %in% 2003:2004) %>% 
  tidytable::mutate(Year = as.factor(YEAR)) -> props.sub

p3.4 = ggplot(data = props.sub, aes(x = rss, colour = Year)) +
  geom_histogram(aes(fill = Year, group = Year), alpha = 0.5, position = "identity", binwidth = 200) +
  theme_classic() +
  theme(legend.position = c(0.8, 0.8),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) +
  labs(x = "RSS", y = "Count", tag = "D") + 
  geom_point(data = props.sub %>% 
               tidytable::left_join(out_4.2$length$iss %>% 
                                      tidytable::select(YEAR, iss)) %>% 
               tidytable::summarise(x = unique(iss), .by = Year) %>% 
               # tidytable::summarise(x = median(rss), .by = Year) %>% 
               tidytable::mutate(y = c(2,2)), aes(x,y, fill = Year), shape = 25, size = 4, color = "black") +
  scale_y_continuous(limits = c(0, 30), expand = c(0, 0)) +
  # scale_x_continuous(n.breaks = 8)
  scale_x_continuous(breaks = seq(0, 35000, by = 2500),
                     labels = c(0, "", 5000, "", 10000, "", 15000, "", 20000, "", 25000, "", 30000, "", 35000),
                     limits = c(0, 35000),
                     expand = c(0, 0))+
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# # now do RSS with boot all + SS
# out_2.2$length$rss %>% 
#   tidytable::filter(YEAR %in% 2003:2004) %>% 
#   tidytable::mutate(Year = as.factor(YEAR)) -> props.sub
# 
# p3.5 = ggplot(data = props.sub, aes(x = rss, colour = Year)) +
#   geom_histogram(aes(fill = Year, group = Year), alpha = 0.5, position = "identity", binwidth = 200) +
#   theme_classic() +
#   theme(legend.position = c(0.8, 0.8),
#         plot.tag.position = c(0.1, 0.95),
#         plot.tag = element_text(size = 14, face = "bold")) +
#   labs(x = "RSS", y = "Count", tag = "E") + 
#   geom_point(data = props.sub %>% 
#                tidytable::left_join(out_2.2$length$iss %>% 
#                                       tidytable::select(YEAR, iss)) %>% 
#                tidytable::summarise(x = unique(iss), .by = Year) %>% 
#                # tidytable::summarise(x = median(rss), .by = Year) %>% 
#                tidytable::mutate(y = c(9,9)), aes(x,y, fill = Year), shape = 25, size = 4, color = "black") +
#   scale_y_continuous(limits = c(0, 30), expand = c(0, 0)) +
#   # scale_x_continuous(n.breaks = 8)
#   scale_x_continuous(breaks = seq(0, 35000, by = 2500),
#                      labels = c(0, "", 5000, "", 10000, "", 15000, "", 20000, "", 25000, "", 30000, "", 35000),
#                      limits = c(0, 35000),
#                      expand = c(0, 0))
# 
#

# 
# p1 / p3
# p1 / p3 / p3.2

p1 / p3.2 / p3.3 / p3.4 / p3

```

The comparison shows that fish in 2003 were less similar in length compared to 2004. 2003 corresponds to a higher ISS even though the sample rate was the same and the number of hauls and nominal sample size were comparable. More of the fish sampled in 2003 were required for characterizing the underlying multinomial sampling distribution compared to 2004 because more lengths were represented at a higher rate in 2003 (shorter, wider distribution). This is reflected in the RSS (ref figure).

# * Time series Expansion NEW
```{r, fig.width=9, fig.height=5}
out_2.1.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_44.1_ISS_500_iters_2025-03-25.RDS")

year = out_2.1.1$expansion_NULL$length$iss$YEAR

df = tidytable(year = year,
               none = out_2.1.1$expansion_NULL$length$iss$iss,
               medium = out_2.1.1$expansion_haul_numbers$length$iss$iss,
               full = out_2.1.1$expansion_haul_month_numbers$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

# 
# # all years
# ggplot(data = df2, aes(x = year, y = ISS, group = scenario, colour = scenario)) +
#   geom_line() +
#   labs(x = "", y = "ISSB") +
#   geom_point() + 
#   theme_classic() +
#   theme(legend.position = c(0.84, 0.8),
#         axis.title.x = element_blank())

# 1999:2023
df2 %>% 
  filter(year%in%1999:2023) %>% 
ggplot(aes(x = year, y = ISS, group = scenario, colour = scenario)) +
  geom_line() +
  geom_point() +
  scale_shape_manual(values = c(1,2,3)) +
  labs(x = "Year", y = "ISSB", color = "Expansion Complexity") +
  scale_color_manual(labels = c("None", "Medium: haul only", "Full: haul, area, gear, month"), values = c("blue", "red", "black")) +
  geom_point() + 
  theme_classic() + 
  theme(legend.position = c(0.7, 0.9),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14))
```

\newpage

# * Reduce Samples

```{r}
# out_31.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_31.1.TAS_ISS_500_iters_2025-03-02.RDS")
out_31.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_47.1_ISS_500_iters_2025-03-30.RDS")

year = out_31.1$samples_1$data_H_P$length$iss$YEAR

### H_P ###################
df = tidytable(year = year,
               N1_HP = out_31.1$samples_1$data_H_P$length$iss$iss,
               N5_HP = out_31.1$samples_5$data_H_P$length$iss$iss,
               N10_HP = out_31.1$samples_10$data_H_P$length$iss$iss,
               N15_HP = out_31.1$samples_15$data_H_P$length$iss$iss,
               N20_HP = out_31.1$samples_20$data_H_P$length$iss$iss,

               NN_HP = out_1.5$length$iss$iss[out_1.5$length$iss$YEAR%in%1999:2023])

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

scenario_names = c("N1_HP", "N5_HP", "N10_HP", "N15_HP", "N20_HP", "NN_HP")

## 1
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="N20_HP"])/df2$ISS[df2$scenario=="N20_HP"],
                    type = "relative_20",
                    .by = scenario) -> df3.1
## 2
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="NN_HP"])/df2$ISS[df2$scenario=="NN_HP"],
                    type = "relative_N",
                    .by = scenario) -> df3.2

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>%
  tidytable::filter(scenario!="NN_HP") %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -4)) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.a

### H ###################
df = tidytable(year = year,
               N1_H = out_31.1$samples_1$data_H$length$iss$iss,
               N5_H = out_31.1$samples_5$data_H$length$iss$iss,
               N10_H = out_31.1$samples_10$data_H$length$iss$iss,
               N15_H = out_31.1$samples_15$data_H$length$iss$iss,
               N20_H = out_31.1$samples_20$data_H$length$iss$iss,

               NN_H = out_1.5$length$iss$iss[out_1.5$length$iss$YEAR%in%1999:2023])

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

scenario_names = c("N1_H", "N5_H", "N10_H", "N15_H", "N20_H", "NN_H")

## 1
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="N20_H"])/df2$ISS[df2$scenario=="N20_H"],
                    type = "relative_20",
                    .by = scenario) -> df3.1
## 2
df2 %>% 
  tidytable::mutate(ISS.error = 100*(ISS - df2$ISS[df2$scenario=="NN_H"])/df2$ISS[df2$scenario=="NN_H"],
                    type = "relative_N",
                    .by = scenario) -> df3.2

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>%
  tidytable::filter(scenario!="NN_H") %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -3)) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.b

```

\newpage

```{r, fig.height=7}
### plots

p1 = df4.a %>%
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "", y = "ISSB % Difference", tag = "A") +
  theme_classic() +
  theme(legend.position = c(0.7, 0.15), 
        legend.direction = "horizontal", 
        legend.box = "horizontal",
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) + 
  guides(fill = guide_legend(title = "haul and port data",
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  # geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-80, 10))

## for presentation
p.p = df4.a %>%
  filter(type=="relative_N") %>% 
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Adjusted Samples Per Haul", y = "ISSB % Difference", tag = "") +
  theme_classic() +
  theme(legend.position = c(3, 0.15),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.text = element_text(size = 12),
        legend.title = element_text(size = 14)) + 
  guides(fill = guide_legend(title = "relative to baseline")) +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  # geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-80, 10))

p2 = df4.b %>%
  ggplot(aes(x = scenario, y = ISS.error)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Samples per Haul", y = "ISSB % Difference", tag = "B") +
  theme_classic() +
  theme(legend.position = c(0.7, 0.15), 
        legend.direction = "horizontal", 
        legend.box = "horizontal",
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) + 
  guides(fill = guide_legend(title = "haul only data",
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  # geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-80, 10))

p1 + plot_spacer() + p2 + plot_layout(ncol = 1,
                           heights = c(6, -.5, 6))

```

\newpage

for above. "the median percent difference in ISSB increased with increasing sample size using the full data set (A) and only the haul data (B) over the period 1999-2023. There was no appreciable change in the percent difference when only using the haul data. There was also very little difference comparing the percent difference relative to the baseline ISSB or relative to applying the sample reduction routine at a sampling rate of 20. This verifies that the results are comparable for the following scenarios: comparisons to the baseline or comparisons to just using hauls with a sampling rate that never exceeds 20. The most conservative scenario is assuming a future sampling where high sampling rates from port samples are not available and the haul sampling rate is allowed to be below 20 but not exceed it. Comparisons relative to this scenario are the most relevant, conservative, scenario available to evaluate the impact of reducing observer haul sampling rate on composition data precision for future years. If the sampling rate behavior continues on average at least as good as it has done for the period 1999:2023 (which it is scheduled to do), we can expect the loss of precision not to exceed these values. (Assuming fishing trip and haul characteristics are comparable on average, also likely expected).

\newpage

# * Samples VS Expantion. intra-relative

Make sure to say in the paper: Methods: Unlike changing the number of hauls or trips (static quality of the data), the stock assessor has discression over changing the expansion method. IF they change it, they will change it for past and present data, not change it based on blocks of time or something like that. SO, these scenarios demonstrate what we can expect to see in near future years IF we decrease the sample size, ASSUMING the expansion complexity will change. For example, IF the expansion complexity will change, what will be the expected loss of precision if we reduce the sample size from 20 to 10. REMEMBER, these scenarios are relative to the N=20 case for each expansion complexity case.

```{r, fig.width=8}

out_22.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_38.2_ISS_500_iters_2025-03-20.RDS")

year = out_22.1$samples_1$expansion_NULL$length$iss$YEAR

df = tidytable(year = year,
               N1_Enone = out_22.1$samples_1$expansion_NULL$length$iss$iss,
               N5_Enone = out_22.1$samples_5$expansion_NULL$length$iss$iss,
               N10_Enone = out_22.1$samples_10$expansion_NULL$length$iss$iss,
               N15_Enone = out_22.1$samples_15$expansion_NULL$length$iss$iss,
               N20_Enone = out_22.1$samples_20$expansion_NULL$length$iss$iss,
               N1_Emedium = out_22.1$samples_1$expansion_haul_numbers$length$iss$iss,
               N5_Emedium = out_22.1$samples_5$expansion_haul_numbers$length$iss$iss,
               N10_Emedium = out_22.1$samples_10$expansion_haul_numbers$length$iss$iss,
               N15_Emedium = out_22.1$samples_15$expansion_haul_numbers$length$iss$iss,
               N20_Emedium = out_22.1$samples_20$expansion_haul_numbers$length$iss$iss,
               N1_Efull = out_22.1$samples_1$expansion_haul_month_numbers$length$iss$iss,
               N5_Efull = out_22.1$samples_5$expansion_haul_month_numbers$length$iss$iss,
               N10_Efull = out_22.1$samples_10$expansion_haul_month_numbers$length$iss$iss,
               N15_Efull = out_22.1$samples_15$expansion_haul_month_numbers$length$iss$iss,
               N20_Efull = out_22.1$samples_20$expansion_haul_month_numbers$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

df2[grep("Enone", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_Enone"])/df2$ISS[df2$scenario=="N20_Enone"],
                    type = "Enone",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -7)) -> df3.1

df2[grep("Emedium", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_Emedium"])/df2$ISS[df2$scenario=="N20_Emedium"],
                    type = "Emedium",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -9)) -> df3.2

df2[grep("Efull", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_Efull"])/df2$ISS[df2$scenario=="N20_Efull"],
                    type = "Efull",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -7)) -> df3.3

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>% 
  tidytable::bind_rows(df3.3) %>% 
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.0

df4 %>% 
  ggplot(aes(x = scenario, y = ISS.error*100)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Samples per Haul", y = "ISSB % Difference") +
  theme_classic() +
  theme(legend.position = c(0.5, 0.2)) +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10)) +
  scale_x_discrete(labels = c("1", "5", "10", "15", "20")) +
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# Stats
df4 %>% filter(type=="trip", scenario=="N_0.25") %>% pull(ISS.error) %>% median
df4 %>% filter(type=="trip", scenario=="N_2.00") %>% pull(ISS.error) %>% median
df4 %>% filter(type=="haul", scenario=="N_0.25") %>% pull(ISS.error) %>% median
df4 %>% filter(type=="haul", scenario=="N_2.00") %>% pull(ISS.error) %>% median

```

# * Haul and TRIP Decrease w/o replacement
```{r}
out_29.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_45.1_ISS_500_iters_2025-03-26.RDS")
out_48.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_48.1_ISS_500_iters_2025-03-27.RDS")
relative.trip = out_48.1$proportion_1$type_trip$length$iss$iss
relative.haul = out_48.1$proportion_1$type_haul$length$iss$iss

year = out_29.1$proportion_0.1$type_trip$length$iss$YEAR

# Trip
df = tidytable(year = year,
               N_0.1 = out_29.1$proportion_0.1$type_trip$length$iss$iss,
               N_0.2 = out_29.1$proportion_0.2$type_trip$length$iss$iss,
               N_0.3 = out_29.1$proportion_0.3$type_trip$length$iss$iss,
               N_0.4 = out_29.1$proportion_0.4$type_trip$length$iss$iss,
               N_0.5 = out_29.1$proportion_0.5$type_trip$length$iss$iss,
               N_0.6 = out_29.1$proportion_0.6$type_trip$length$iss$iss,
               N_0.7 = out_29.1$proportion_0.7$type_trip$length$iss$iss,
               N_0.8 = out_29.1$proportion_0.8$type_trip$length$iss$iss,
               N_0.9 = out_29.1$proportion_0.9$type_trip$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")
# 
# df2 %>% 
#   tidytable::mutate(ISS.error = (ISS - out_1.5$length$iss$iss)/out_1.5$length$iss$iss,
#                     type = "trip",
#                     .by = scenario) -> df3.1
df2 %>% 
  tidytable::mutate(ISS.error = (ISS - relative.trip)/relative.trip,
                    type = "trip",
                    .by = scenario) -> df3.1

# Haul
df = tidytable(year = year,
               N_0.1 = out_29.1$proportion_0.1$type_haul$length$iss$iss,
               N_0.2 = out_29.1$proportion_0.2$type_haul$length$iss$iss,
               N_0.3 = out_29.1$proportion_0.3$type_haul$length$iss$iss,
               N_0.4 = out_29.1$proportion_0.4$type_haul$length$iss$iss,
               N_0.5 = out_29.1$proportion_0.5$type_haul$length$iss$iss,
               N_0.6 = out_29.1$proportion_0.6$type_haul$length$iss$iss,
               N_0.7 = out_29.1$proportion_0.7$type_haul$length$iss$iss,
               N_0.8 = out_29.1$proportion_0.8$type_haul$length$iss$iss,
               N_0.9 = out_29.1$proportion_0.9$type_haul$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

# 
# df2 %>% 
#   tidytable::mutate(ISS.error = (ISS - out_1.5$length$iss$iss)/out_1.5$length$iss$iss,
#                     type = "haul",
#                     .by = scenario) -> df3.2
df2 %>% 
  tidytable::mutate(ISS.error = (ISS - relative.haul)/relative.haul,
                    type = "haul",
                    .by = scenario) -> df3.2

df3.1 %>% 
  tidytable::bind_rows(df3.2) -> df4

df4 %>% 
  ggplot(aes(x = scenario, y = ISS.error*100)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "Proportional Change", y = "ISSB % Difference") +
  theme_classic() +
  theme(legend.position = c(0.8, 0.2)) +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  scale_y_continuous(breaks = seq(-100, 20, by = 10), limits = c(-100, 20)) +
  scale_x_discrete(labels = c("0.1", "0.2", "0.3", "0.4", "0.5", "0.6", "0.7", "0.8", "0.9"))

```

# Samples VS Haul/Trip

### DATA: Samples VS Haul. Replace=TRUE, including 1.0
```{r, fig.width=8}

out_24.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_40.3_ISS_500_iters_2025-03-25.RDS")

year = out_24.1$samples_1$haul_0.25$length$iss$YEAR

df = tidytable(year = year,
               N1_H0.25 = out_24.1$samples_1$haul_0.25$length$iss$iss,
               N5_H0.25 = out_24.1$samples_5$haul_0.25$length$iss$iss,
               N10_H0.25 = out_24.1$samples_10$haul_0.25$length$iss$iss,
               N15_H0.25 = out_24.1$samples_15$haul_0.25$length$iss$iss,
               N20_H0.25 = out_24.1$samples_20$haul_0.25$length$iss$iss,
               N1_H0.5 = out_24.1$samples_1$haul_0.5$length$iss$iss,
               N5_H0.5 = out_24.1$samples_5$haul_0.5$length$iss$iss,
               N10_H0.5 = out_24.1$samples_10$haul_0.5$length$iss$iss,
               N15_H0.5 = out_24.1$samples_15$haul_0.5$length$iss$iss,
               N20_H0.5 = out_24.1$samples_20$haul_0.5$length$iss$iss,
               N1_H0.75 = out_24.1$samples_1$haul_0.75$length$iss$iss,
               N5_H0.75 = out_24.1$samples_5$haul_0.75$length$iss$iss,
               N10_H0.75 = out_24.1$samples_10$haul_0.75$length$iss$iss,
               N15_H0.75 = out_24.1$samples_15$haul_0.75$length$iss$iss,
               N20_H0.75 = out_24.1$samples_20$haul_0.75$length$iss$iss,
               N1_H1.0 = out_24.1$samples_1$haul_1$length$iss$iss,
               N5_H1.0 = out_24.1$samples_5$haul_1$length$iss$iss,
               N10_H1.0 = out_24.1$samples_10$haul_1$length$iss$iss,
               N15_H1.0 = out_24.1$samples_15$haul_1$length$iss$iss,
               N20_H1.0 = out_24.1$samples_20$haul_1$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

df2[grep("H0.25", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_H1.0"])/df2$ISS[df2$scenario=="N20_H1.0"],
                    type = "0.25",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -7)) -> df3.1

df2[grep("H0.5", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_H1.0"])/df2$ISS[df2$scenario=="N20_H1.0"],
                    type = "0.5",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -6)) -> df3.2

df2[grep("H0.75", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_H1.0"])/df2$ISS[df2$scenario=="N20_H1.0"],
                    type = "0.75",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -7)) -> df3.3

df2[grep("H1.0", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_H1.0"])/df2$ISS[df2$scenario=="N20_H1.0"],
                    type = "1.0",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -6)) -> df3.4

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>% 
  tidytable::bind_rows(df3.3) %>% 
  tidytable::bind_rows(df3.4) %>% 
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.1.b

```

### DATA: Samples VS Trip. Replace=TRUE, including 1.0
```{r, fig.width=8}

out_23.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_39.7_ISS_500_iters_2025-03-25.RDS")
# out_23.1 = readRDS(file = "C:/Users/bstacy2/OneDrive - UW/UW Postdoc/GitHub Repos/baseISS_data/outputs/paper_output/paper_scenario_39.8_ISS_500_iters_2025-03-27.RDS")

year = out_23.1$samples_1$trip_0.25$length$iss$YEAR

df = tidytable(year = year,
               N1_T0.25 = out_23.1$samples_1$trip_0.25$length$iss$iss,
               N5_T0.25 = out_23.1$samples_5$trip_0.25$length$iss$iss,
               N10_T0.25 = out_23.1$samples_10$trip_0.25$length$iss$iss,
               N15_T0.25 = out_23.1$samples_15$trip_0.25$length$iss$iss,
               N20_T0.25 = out_23.1$samples_20$trip_0.25$length$iss$iss,
               N1_T0.5 = out_23.1$samples_1$trip_0.5$length$iss$iss,
               N5_T0.5 = out_23.1$samples_5$trip_0.5$length$iss$iss,
               N10_T0.5 = out_23.1$samples_10$trip_0.5$length$iss$iss,
               N15_T0.5 = out_23.1$samples_15$trip_0.5$length$iss$iss,
               N20_T0.5 = out_23.1$samples_20$trip_0.5$length$iss$iss,
               N1_T0.75 = out_23.1$samples_1$trip_0.75$length$iss$iss,
               N5_T0.75 = out_23.1$samples_5$trip_0.75$length$iss$iss,
               N10_T0.75 = out_23.1$samples_10$trip_0.75$length$iss$iss,
               N15_T0.75 = out_23.1$samples_15$trip_0.75$length$iss$iss,
               N20_T0.75 = out_23.1$samples_20$trip_0.75$length$iss$iss,
               N1_T1.0 = out_23.1$samples_1$trip_1$length$iss$iss,
               N5_T1.0 = out_23.1$samples_5$trip_1$length$iss$iss,
               N10_T1.0 = out_23.1$samples_10$trip_1$length$iss$iss,
               N15_T1.0 = out_23.1$samples_15$trip_1$length$iss$iss,
               N20_T1.0 = out_23.1$samples_20$trip_1$length$iss$iss)

df2 = melt.data.table(df, id.vars = "year", variable.name = "scenario", value.name = "ISS")

df2[grep("T0.25", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_T1.0"])/df2$ISS[df2$scenario=="N20_T1.0"],
                    type = "0.25",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -7)) -> df3.1

df2[grep("T0.5", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_T1.0"])/df2$ISS[df2$scenario=="N20_T1.0"],
                    type = "0.5",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -6)) -> df3.2

df2[grep("T0.75", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_T1.0"])/df2$ISS[df2$scenario=="N20_T1.0"],
                    type = "0.75",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -7)) -> df3.3

df2[grep("T1.0", scenario)] %>% 
  tidytable::mutate(ISS.error = (ISS - df2$ISS[df2$scenario=="N20_T1.0"])/df2$ISS[df2$scenario=="N20_T1.0"],
                    type = "1.0",
                    .by = scenario) %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 1, end = -6)) -> df3.4

df3.1 %>% 
  tidytable::bind_rows(df3.2) %>% 
  tidytable::bind_rows(df3.3) %>% 
  tidytable::bind_rows(df3.4) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) -> df4
df4 -> df4.1.a

```

### * 2 Panel: Sample Size VS Trip, Haul

```{r, fig.height=12, fig.width=10}
p1 = df4.1.a %>%
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -1)) %>%
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) %>%
  ggplot(aes(x = scenario, y = ISS.error*100)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "", y = "ISSB % Difference", tag = "A") +
  theme_classic() +
  theme(legend.position = c(0.74, 0.15),
        legend.direction = "horizontal",
        legend.box = "horizontal",
        legend.background = element_rect(fill = "transparent"),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) +
  guides(fill = guide_legend(title = "Trips",
                             title.position = "top",
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-94, 10)) +
  scale_color_viridis_d() +
  scale_fill_viridis_d()

p2 = df4.1.b %>% 
  tidytable::mutate(scenario = str_sub(scenario, start = 2, end = -1)) %>% 
  tidytable::mutate(scenario = forcats::fct_inorder(scenario)) %>% 
  ggplot(aes(x = scenario, y = ISS.error*100)) +
  geom_boxplot(aes(fill = type)) +
  labs(x = "", y = "ISSB % Difference", tag = "B") +
  theme_classic() +
  theme(legend.position = c(0.74, 0.15), 
        legend.direction = "horizontal", 
        legend.box = "horizontal",
        legend.background = element_rect(fill = "transparent"),
        plot.tag.position = c(0.1, 0.95),
        plot.tag = element_text(size = 14, face = "bold")) + 
  guides(fill = guide_legend(title = "Hauls",
                             title.position = "top", 
                             title.hjust = 0.5,
                             label.position = "bottom"))  +
  geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
  geom_hline(yintercept = 0, linetype = "solid", color = "black") +
  scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-94, 10)) +
  scale_color_viridis_d() +
  scale_fill_viridis_d()

# p3 = df4.3.b %>% 
#   tidytable::mutate(scenario = factor(str_sub(scenario, start = 2, end = -1))) %>% 
#   tidytable::filter(type %in% scenario_names.short[c(2,4,5,6,7)]) %>%
#   tidytable::mutate(scenario = forcats::fct_inorder(scenario)) %>% 
#   ggplot(aes(x = scenario, y = ISS.error*100)) +
#   geom_boxplot(aes(fill = type)) +
#   labs(x = "Samples per Haul", y = "ISSB % Difference", tag = "C") +
#   theme_classic() +
#   theme(legend.position = c(0.7, 0.15), 
#         legend.direction = "horizontal", 
#         legend.box = "horizontal",
#         plot.tag.position = c(0.1, 0.95),
#         plot.tag = element_text(size = 14, face = "bold")) + 
#   guides(fill = guide_legend(title = "Hauls/Trips",
#                              title.position = "top", 
#                              title.hjust = 0.5,
#                              label.position = "bottom"))  +
#   geom_hline(yintercept = -10, linetype = "dashed", color = "black") +
#   geom_hline(yintercept = 10, linetype = "dashed", color = "black") +
#   scale_y_continuous(breaks = seq(-100, 120, by = 10), limits = c(-90, 30))

p1 + plot_spacer() + p2 + plot_layout(ncol = 1,
                           heights = c(6, -.5, 6))

```

# SCRATCH
## PLOT BOOT.HAUL AND BOOT.TRIP ONLY SCENARIOS?

```{r}

par(mfrow = c(1,1))
plot(out_37.2$length$iss$YEAR, out_37.2$length$iss$iss, col = "blue", type = "l")
lines(out_1.5$length$iss$YEAR, out_1.5$length$iss$iss)
lines(out_37.3$length$iss$YEAR, out_37.3$length$iss$iss, col = "purple")

```